---
title: "Simple document"
output: github_document
---
 
```{r}
library(tidyverse)
library(readxl)
library(haven)
```
  
  
Let's import the `FAS_litters.csv` csv using a relative path. 
  
```{r}
litters_df =
  read_csv("data/FAS_litters.csv")

litters_df =
  janitor::clean_names(litters_df)
``` 

Import the same dataset using an absolute path. 

```{r, eval = FALSE}
litters_df_abs =
  read_csv("~/Desktop/Fall 2023/Data Science I/building blocks R studio/data_wrangling_i/data/FAS_litters.csv")
 
litters_df_abs =
  janitor::clean_names(litters_df_abs)
``` 

Learning Assessment: loading the `FAS_pups.csv` dataset 
  Using an relative path: 
```{r}
pups_df =
  read_csv("data/FAS_pups.csv")

pups_df =
  janitor::clean_names(pups_df)
```   
 
 Absolute Path: 
```{r, eval = FALSE}
pups_df_abs =
  read_csv("~/Desktop/Fall 2023/Data Science I/building blocks R studio/data_wrangling_i/data/FAS_pups.csv")
 
pups_df_abs =
  janitor::clean_names(pups_df_abs)
``` 

## look at the data
```{r}
litters_df

head(litters_df)

tail(litters_df)
``` 

You cab use `view` to see the entire dataset
```{r, eval = FALSE}
litters_df
view(litters_df)
``` 

Look at a data summary
```{r, eval = FALSE}
str(litters_df)

skimr::skim(litters_df)
``` 

## Options in `read_*`
```{r}
litters_df =
  read_csv("data/FAS_litters.csv", 
           skip = 10, col_names = FALSE)
``` 

Look at NA values.
```{r}
litters_df =
  read_csv("data/FAS_litters.csv", 
           na = c("NA", 19))
``` 

Column types
```{r}
litters_df =
  read_csv("data/FAS_litters.csv", 
           col_types =
             cols(
               `GD0 weight` = col_character()
               
             ))
``` 

Import a xlsx file. 
```{r}
mlb11_df = 
  read_excel("data/mlb11.xlsx")

``` 

Import a SAS file.

```{r}
pulse_df=
  read_sas("data/public_pulse_data.sas7bdat")
```

## Base R ...
```{r, eval = FALSE}
litters_df=
  read.csv("data/FAS_litters.csv")
```


## Export data

We have a code that "cleans" data and need to export the result. 

```{r}
litters_df_cleaned=
  read_csv("data/FAS_litters.csv")

litters_df_cleaned=
  janitor::clean_names(litters_df_cleaned)

write_csv(litters_df_cleaned, "data/litters_cleaned.csv")
```


